---
title: "ICY0006: Statistics and probabilty theory project"
output: html_notebook
---
Markus Aas 192920IVSB


### Stage 1:
My chosen dataset is about [Apartment prices in Sao Paulo](https://www.kaggle.com/felipecabueno/apartment-prices-in-sao-paulo-br-2020), which I got from the website [Kaggle](https://www.kaggle.com/).

The dataset has 7 variables in total, of which 6 are numerical and 1 is categorical. The categorical variable in our case is

* neighborhood name

and the numerical variables are

* area of the apartment in square meters 
* number of bedrooms 
* number of bathrooms
* number of suites (set of rooms, usually bedroom with its own bathroom)
* number of parking spaces
* final price of the apartment in thousands of reais (R$, Brazilian currency).

Here is an example of the 10 first rows of my dataset and basic information about the dataset:
```{r}
apartments_sp <- read.csv("C:/Users/aasma/Statistics/apartments_sp.csv")
head(apartments_sp, n=10)
str(apartments_sp)
```

The dataset has no missing values or errors.

Some data visualization through distribution graphs:
```{r}
hist(apartments_sp$square_meters, main="Histogram for the area of the apartments", xlab="Square meters", col="red", breaks=100)
hist(apartments_sp$bedrooms, breaks=8, main="Histogram for the number of bedrooms", xlab="Number of bedrooms")
hist(apartments_sp$bathrooms, breaks=10, main="Histogram for the number of bathrooms", xlab="Number of bathrooms")
hist(apartments_sp$suites, breaks=8, main="Histogram for the number of suites", xlab="Number of suites")
hist(apartments_sp$parking_spaces, breaks=10, main="Histogram for the number of parking spaces", xlab="Number of parking spaces")
hist(apartments_sp$final_price, main="Histogram for apartment prices", xlab="Thousands of reais (R$)", col="red", breaks=100)
```
```{r}
lines(apartments_sp$square_meters)
```

