---
title: "ICY0006: Statistics and probabilty theory project"
output: html_notebook
---
Markus Aas 192920IVSB


### Stage 1: Information about the dataset and distribution graphs.
My chosen dataset is about [Apartment prices in Sao Paulo](https://www.kaggle.com/felipecabueno/apartment-prices-in-sao-paulo-br-2020), which I got from the website [Kaggle](https://www.kaggle.com/).

The dataset has 7 variables in total, of which 6 are numerical and 1 is categorical. The categorical variable in our case is

* neighborhood name

and the numerical variables are

* area of the apartment in square meters 
* number of bedrooms 
* number of bathrooms
* number of suites (set of rooms, usually bedroom with its own bathroom)
* number of parking spaces
* final price of the apartment in thousands of reais (R$, Brazilian currency).

Here is an example of the 10 first rows of my dataset and basic information about the dataset:
```{r}
apartments_sp <- read.csv("C:/Users/aasma/Statistics/apartments_sp.csv")
head(apartments_sp, n=10)
str(apartments_sp)
```

The dataset has no missing values or errors.

The following histograms and kernel density plots give an approximate understanding of the variables in the dataset and showcase the frequency and values of those variables. 
```{r}
hist(apartments_sp$square_meters, main="Histogram of the area of the apartments", xlab="Square meters", col="red", breaks=100)
hist(apartments_sp$bedrooms, breaks=8, main="Histogram of the number of bedrooms", xlab="Number of bedrooms")
hist(apartments_sp$bathrooms, breaks=10, main="Histogram of the number of bathrooms", xlab="Number of bathrooms")
hist(apartments_sp$suites, breaks=8, main="Histogram of the number of suites", xlab="Number of suites")
hist(apartments_sp$parking_spaces, breaks=10, main="Histogram of the number of parking spaces", xlab="Number of parking spaces")
hist(apartments_sp$final_price, main="Histogram of apartment prices", xlab="Thousands of reais (R$)", col="red", breaks=100)
```
```{r}
plot(density(apartments_sp$square_meters), main = "Kernel density of area of the apartments", xlab="Square meters")
plot(density(apartments_sp$bedrooms), main = "Kernel density of the number of bedrooms", xlab="Number of bedrooms")
plot(density(apartments_sp$bathrooms), main = "Kernel density of the number of bathrooms", xlab="Number of bathrooms")
plot(density(apartments_sp$suites), main = "Kernel density of the number of suites", xlab="Number of suites")
plot(density(apartments_sp$parking_spaces), main = "Kernel density of the number of parking spaces", xlab="Number of parking spaces")
plot(density(apartments_sp$final_price), main = "Kernel density of apartment prices", xlab="Thousands of reais (R$)")
```

### Stage 2: Central tendency and variability

```{r}
mean(apartments_sp$square_meters)
```



