---
title: "ICY0006: Statistics and probabilty theory project"
output: html_notebook
---
Markus Aas 192920IVSB


### Stage 1: Information about the dataset and distribution graphs.
My chosen dataset is about [Apartment prices in Sao Paulo](https://www.kaggle.com/felipecabueno/apartment-prices-in-sao-paulo-br-2020), which I got from the website [Kaggle](https://www.kaggle.com/).

The dataset has 7 variables in total, of which 6 are numerical and 1 is categorical. The categorical variable in our case is

* neighborhood name

and the numerical variables are

* area of the apartment in square meters 
* number of bedrooms 
* number of bathrooms
* number of suites (set of rooms, usually bedroom with its own bathroom)
* number of parking spaces
* final price of the apartment in thousands of reais (R$, Brazilian currency).

Here is an example of the 10 first rows of my dataset and basic information about the dataset:
```{r}
apartments_sp <- read.csv("C:/Users/aasma/Statistics/apartments_sp.csv")
head(apartments_sp, n=10)
str(apartments_sp)
```

The dataset has no missing values or errors.

The following histograms and kernel density plots give an approximate understanding of the variables in the dataset and showcase the frequency and values of those variables. 
```{r}
library(ggplot2)

ggplot(apartments_sp, aes(x=square_meters)) +
  geom_histogram(binwidth=10) +
  labs(title="Histogram of the area of the apartments", x="Square meters")

ggplot(apartments_sp, aes(x=bedrooms)) +
  geom_histogram(binwidth=1) +
  labs(title="Histogram of the number of bedrooms", x="Number of bedrooms")

ggplot(apartments_sp, aes(x=bathrooms)) +
  geom_histogram(binwidth=1) +
  labs(title="Histogram of the number of bathrooms", x="Number of bathrooms")

ggplot(apartments_sp, aes(x=suites)) +
  geom_histogram(binwidth=1) +
  labs(title="Histogram of the number of suites", x="Number of suites")

ggplot(apartments_sp, aes(x=parking_spaces)) +
  geom_histogram(binwidth=1) +
  labs(title="Histogram of the number of parking spaces", x="Number of parking spaces")

ggplot(apartments_sp, aes(x=final_price)) +
  geom_histogram(binwidth=100) +
  labs(title="Histogram of apartment prices", x="Thousands of reais (R$)")
```
```{r}
library(ggplot2)

ggplot(apartments_sp, aes(x=square_meters)) +
  geom_freqpoly(size=0.8) +
  labs(title="Frequency polygon of area of the apartments", x="Square meters")

ggplot(apartments_sp, aes(x=bedrooms)) +
  geom_freqpoly(size=0.8) +
  labs(title="Frequency polygon of the number of bedrooms", x="Number of bedrooms")

ggplot(apartments_sp, aes(x=bathrooms)) +
  geom_freqpoly(size=0.8) +
  labs(title="Frequency polygon of the number of bathrooms", x="Number of bathrooms")

ggplot(apartments_sp, aes(x=suites)) +
  geom_freqpoly(size=0.8) +
  labs(title="Frequency polygon of the number of suites", x="Number of suites")

ggplot(apartments_sp, aes(x=parking_spaces)) +
  geom_freqpoly(size=0.8) +
  labs(title="Frequency polygon of the number of parking spaces", x="Number of parking spaces")

ggplot(apartments_sp, aes(x=final_price)) +
  geom_freqpoly(size=0.8) +
  labs(title="Frequency polygon of apartment prices", x="Thousands of reais (R$)")

```

### Stage 2: Central tendency and variability

```{r}
library(knitr)
df <- data.frame(
  Variable = c("Square meters", "Bedrooms", "Bathrooms", "Suites", "Parking spaces", "Final price"),
  Mean = c(mean(apartments_sp$square_meters), mean(apartments_sp$bedrooms), mean(apartments_sp$bathrooms), mean(apartments_sp$suites),
           mean(apartments_sp$parking_spaces), mean(apartments_sp$final_price)),
  Median = c(median(apartments_sp$square_meters), median(apartments_sp$bedrooms), median(apartments_sp$bathrooms), median(apartments_sp$suites),
             median(apartments_sp$parking_spaces), median(apartments_sp$final_price)),
  Mode = c(names(table(apartments_sp$square_meters))[table(apartments_sp$square_meters)==max(table(apartments_sp$square_meters))],
           names(table(apartments_sp$bedrooms))[table(apartments_sp$bedrooms)==max(table(apartments_sp$bedrooms))],
           names(table(apartments_sp$bathrooms))[table(apartments_sp$bathrooms)==max(table(apartments_sp$bathrooms))],
           names(table(apartments_sp$suites))[table(apartments_sp$suites)==max(table(apartments_sp$suites))],
           names(table(apartments_sp$parking_spaces))[table(apartments_sp$parking_spaces)==max(table(apartments_sp$parking_spaces))],
           names(table(apartments_sp$final_price))[table(apartments_sp$final_price)==max(table(apartments_sp$final_price))])
)
kable(df, caption="Central tendency table")
```

```{r}
library(knitr)
df1 <- data.frame(
   Variable = c("Square meters", "Bedrooms", "Bathrooms", "Suites", "Parking spaces", "Final price"),
   Range = c(max(apartments_sp$square_meters) - min(apartments_sp$square_meters), max(apartments_sp$bedrooms) - min(apartments_sp$bedrooms),
             max(apartments_sp$bathrooms) - min(apartments_sp$bathrooms), max(apartments_sp$suites) - min(apartments_sp$suites),
             max(apartments_sp$parking_spaces) - min(apartments_sp$parking_spaces), max(apartments_sp$final_price) - min(apartments_sp$final_price)),
   Interquartile_range = c(IQR(apartments_sp$square_meters), IQR(apartments_sp$bedrooms), IQR(apartments_sp$bathrooms),
                           IQR(apartments_sp$suites), IQR(apartments_sp$parking_spaces), IQR(apartments_sp$final_price)),
   Standard_deviation = c(sd(apartments_sp$square_meters), sd(apartments_sp$bedrooms), sd(apartments_sp$bathrooms),
                          sd(apartments_sp$suites), sd(apartments_sp$parking_spaces), sd(apartments_sp$final_price)),
   Variance = c("-", var(apartments_sp$bedrooms), var(apartments_sp$bathrooms), var(apartments_sp$suites), var(apartments_sp$parking_spaces), "-")
)
kable(df1, caption="Variability table")

```


### Stage 3:

```{r}
library(ggplot2)
library(knitr)

df2 <- apartments_sp[ , 2:7]

kable(cor(df2))

```

As we can see from the correlation matrix, all of the numerical values have linear relationships with each other from moderate to strong correlation.
So here are some scatter plots of variables with the highest correlation coefficients.

```{r}
library(ggplot2)

ggplot(apartments_sp, aes(x=square_meters, y=final_price)) +
  geom_point()

ggplot(apartments_sp, aes(x=bathrooms, y=suites)) +
  geom_jitter()

ggplot(apartments_sp, aes(x=square_meters, y=bathrooms)) +
  geom_jitter()

ggplot(apartments_sp, aes(x=bathrooms, y=parking_spaces)) +
  geom_jitter()

ggplot(apartments_sp, aes(x=square_meters, y=parking_spaces)) +
  geom_jitter()


```

For step 3 I had to choose the pair square_meters vs final_price, because these are the only continuous variables in my dataset, meaning I can only use these variables for linear regression.

```{r}

lmMod <- lm(apartments_sp$final_price ~ apartments_sp$square_meters)
lmMod

ggplot(apartments_sp, aes(x=square_meters, y=final_price)) +
  geom_point() +
  geom_smooth(method=lm)

```

From the coefficients in the previous output we can construct our linear regression equation which in our case will be 
final_price = 59.908 + 7.308 * square_meters.


### Stage 4:

For stage 4 I chose the lottery EuroMillions.
Players have to choose 5 main numbers from 1 to 50 and 2 extra lucky star numbers from 1 to 12.

To compute the probabilites to win the largest prize, we are going to use the following formula on both the main numbers and lucky star numbers:
n! / (r! * (n - r)!)
where n is the total number of possible number and r is the number of numbers chosen.
So the computation in our case for main prize (5 main numbers guessed and 2 lucky star numbers guessed) would be as following:

```{r}
main_numbers <- factorial(50) / (factorial(5) * factorial(50 - 5))
lucky_stars <- factorial(12) / (factorial(2) * factorial(12 - 2))

main_numbers * lucky_stars

```
We have 1 in 139 838 160 chance of winning the main prize.

For odds for smaller winnings we have to use a more complicated formula:
(r! / (k! * (r - k)!)) * ((n-r)! / ((n - r) - (r - k))! * (r - k)!)
where k is the numbers chosen correctly.

So now that we have all the necessary formulas, we can compute the rest of the winning chances.
For simplicity I am not going to write the full formulas if computation are easy.

The computation for 5 main numbers and 1 lucky star number guessed:
```{r}
one_lucky_star <- lucky_stars / (2 * (factorial(10) / factorial(9)))

main_numbers * one_lucky_star
```
 The probability is 1 in 6 991 908.
 
 For 5 main numbers and  0 lucky star numbers:
```{r}
zero_lucky_starts <- lucky_stars / (factorial(10) / (factorial(8) * 2))

main_numbers * zero_lucky_starts
```
 The probability is 1 in 3 107 515.
 
 For 4 main numbers and 2 lucky star numbers:
```{r}
four_main_numbers <- main_numbers / ((factorial(5) / factorial(4)) * (factorial(45) / factorial(44)))
four_main_numbers * lucky_stars
```
 The probability is 1 in 621 503.
 
 For 4 main numbers and 1 lucky star number:
```{r}
four_main_numbers * one_lucky_star
```
 The probability is 1 in 31 075.
 
 For 3 main numbers and 2 lucky star numbers:
```{r}
three_main_numbers <- main_numbers / ((factorial(5) / 12) * (factorial(45) / (factorial(43) * 2)))
three_main_numbers * lucky_stars
```
The probability is 1 in 14 125.

For 4 main numbers and 0 lucky star numbers:
```{r}
four_main_numbers * zero_lucky_starts
```
The probability is 1 in 13811.

For 2 main number and 2 lucky star numbers:
```{r}
two_main_numbers <- main_numbers / ((factorial(5) / 12) * (factorial(45) / (factorial(42) * 6)))
two_main_numbers * lucky_stars
```
The probability is 1 in 985.

For 3 main numbers and 1 lucky star number:
```{r}
three_main_numbers * one_lucky_star
```
The probability is 1 in 706.

For 3 main numbers and 0 lucky star numbers:
```{r}
three_main_numbers * zero_lucky_starts
```
The probability is 1 in 314.

For 1 main number and 2 lucky star numbers:
```{r}
one_main_number <- main_numbers / ((factorial(5) / 24) * (factorial(45) / (factorial(41) * 24)))
one_main_number * lucky_stars
```
The probability is 1 in 188.

For 2 main numbers and 1 lucky star number:
```{r}
two_main_numbers * one_lucky_star
```
The probability is 1 in 49.

For 2 main numbers and 0 lucky star numbers:
```{r}
two_main_numbers * zero_lucky_starts
```
The probability is 1 in 22.

[EuroMillions wikipedia](https://en.wikipedia.org/wiki/EuroMillions)
[Formula source](https://www.wikihow.com/Calculate-Lotto-Odds)


### Stage 5:






